---
title: 'Reproducible Research: Peer Assessment 1'
author: "Mauricio"
output: html_document
---


## Loading and preprocessing the data

```{r libraries}
library(dplyr)
library(ggplot2)
```

```{r data}
activity <- read.csv("activity.csv")
activity$steps <- as.numeric(activity$steps)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

Total steps per day:

```{r stepsperday}
grp_steps_day <- group_by(activity,date)
total_steps <- summarise(grp_steps_day, total_steps = sum(steps,na.rm = TRUE))
qplot(total_steps$total_steps, geom = "histogram",xlab="# Steps per day", ylab="Frequency", binwidth=500)
stepsmean <- mean(total_steps$total_steps)
stepsmedian <- median(total_steps$total_steps)
```


- Mean = `r stepsmean`
- Median = `r stepsmedian`

## What is the average daily activity pattern?

Average by interval:

```{r averageinterval}
grp_steps_interval <- group_by(activity,interval)
average_steps <- summarise(grp_steps_interval, average_steps = mean(steps,na.rm = TRUE))
qplot(x = average_steps$interval, y = average_steps$average_steps,geom = "line",xlab="5-minute interval", ylab="Mean Steps")
max<-max(average_steps$average_steps)
intervalmax <- average_steps[which(average_steps$average_steps==max),1]
```

The interval `r intervalmax` contain the maximum average steps, which is `r max`

## Imputing missing values

```{r na}
total_na <- sum(is.na(activity$steps))
activity_nNA <- activity

for (i in 1:dim(activity_nNA)[1]){
  if(is.na(activity_nNA[i,1])){
    activity_nNA[i,1] <- average_steps[activity_nNA[i,3]==average_steps$interval,2]
  }
}

grp_steps_day2 <- group_by(activity_nNA,date)
total_steps2 <- summarise(grp_steps_day2, total_steps = sum(steps,na.rm = TRUE))
qplot(total_steps2$total_steps, geom = "histogram",xlab="# Steps per day", ylab="Frequency", binwidth=500)
stepsmean2 <- mean(total_steps2$total_steps)
stepsmedian2 <- median(total_steps2$total_steps)
```

- Total NAs = `r total_na`
- Mean without NA = `r stepsmean2`
- Median without NA = `r stepsmedian2`

As expected, adding values to the NA rows in the dateset has changed the mean and median observed. Depeding on the purpose of these calculations, these impacts can alter the conclusions taken in the experiment.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays}
activity_nNA$weekday <- ifelse(weekdays(activity_nNA$date)%in%c("sábado","domingo"),"weekend","weekday")
activity_nNA$weekday <- as.factor(activity_nNA$weekday)

grp_steps_interval2 <- group_by(activity_nNA,interval,weekday)
average_steps2 <- summarise(grp_steps_interval2, average_steps = mean(steps,na.rm = TRUE))
ggplot(average_steps2, aes(interval, average_steps)) + geom_line() + facet_grid (weekday~.)+
    xlab("5-minute Interval") + 
    ylab("Avarage Steps")
```
